Connecting the Popularity Adjusted Block Model to the Generalized Random Dot Product Graph for Community Detection and Parameter Estimation

### 2021-06-09 summary.Rmd

\citet*{noroozi2019estimation} used Sparse Subspace Clustering (SSC) for 
community detection in the PABM. 
SSC is performed by solving an optimization problem for each observed point.
Given $X \in \mathbb{R}^{n \times d}$ with vectors 
$x_i^\top \in \mathbb{R}^d$ as rows of $X$, the optimization problem 
$\min_{c_i} ||c_i||_1$ subject to $x_i = X c_i$ and $\beta_i = 0$ is solved 
for each $i = 1, ..., n$. The solutions are collected collected into matrix 
$C = \begin{bmatrix} c_1 & \cdots & c_n \end{bmatrix}^\top$ to construct an 
affinity matrix $B = |C| + |C^\top|$. If each $x_i$ lie perfectly on one of $K$ 
subspaces, $B$ describes an undirected graph consisting of $K$ disjoint 
subgraphs, i.e., $B_{ij} = 0$ if $x_i, x_j$ are in different subspaces. 
If $X$ instead represents points near $K$ subspaces with some noise, a final 
graph partitioning step is performed (e.g., edge thresholding or spectral 
clustering).

@@

The explicit form of $XU$ represents points in 
$\mathbb{R}^{K^2}$ such that points within each community lie on 
$K$-dimensional orthogonal subspaces. 
Multiplication by $Q \in \mathbb{O}(p, q)$ removes the orthogonality property 
but retains the property that each community is represented by a 
$K$-dimensional subspace. Therefore, the ASE of $P$ results in subspaces that 
correspond to each community, suggesting the use of SSC. In this paper, we will 
use a different and leave the properties of SSC on the ASE of $A$ as future
work.

@@

\begin{theorem}
\label{theorem5}
Let $P_n$ describe the edge probability matrix of the PABM with 
$n$ vertices, and let $A_n \sim \Bernoulli(P_n)$.  Let $\hat{V}_n$ be the 
matrix of eigenvectors of $A_n$ corresponding to the $K (K + 1) / 2$ most
positive and $K (K - 1) / 2$ most negative eigenvalues. Then 
$\exists \lambda > 0$ such that as $n \to \infty$, $\hat{V}_n$ obeys the 
subspace detection property with probability 1.
\end{theorem}

>>>>

\citet*{noroozi2019estimation} used Sparse Subspace Clustering (SSC) for 
community detection in the PABM. 
SSC is performed by solving an optimization problem for each observed point.
Given $X \in \mathbb{R}^{n \times d}$ with vectors 
$x_i^\top \in \mathbb{R}^d$ as rows of $X$, the optimization problem 
$c_i = \arg\min_{c} ||c||_1$ subject to $x_i = X c$ and $c^{(i)} = 0$ is solved 
for each $i = 1, ..., n$. The solutions are collected collected into matrix 
$C = \begin{bmatrix} c_1 & \cdots & c_n \end{bmatrix}^\top$ to construct an 
affinity matrix $B = |C| + |C^\top|$. If each $x_i$ lie perfectly on one of $K$ 
subspaces, $B$ describes an undirected graph consisting of $K$ disjoint 
subgraphs, i.e., $B_{ij} = 0$ if $x_i, x_j$ are in different subspaces. 
If $X$ instead represents points near $K$ subspaces with some noise, a final 
graph partitioning step may be performed (e.g., edge thresholding or spectral 
clustering).

@@

The explicit form of $XU$ represents points in 
$\mathbb{R}^{K^2}$ such that points within each community lie on 
$K$-dimensional orthogonal subspaces. 
Multiplication by $Q \in \mathbb{O}(p, q)$ removes the orthogonality property 
but retains the property that each community is represented by a 
$K$-dimensional subspace. Therefore, the ASE of $P$ results in subspaces that 
correspond to each community, suggesting the use of SSC. Before exploring 
SSC, we will first consider a different approach.

@@

\begin{theorem}
\label{theorem5}
Let $P_n$ describe the edge probability matrix of the PABM with 
$n$ vertices, and let $A_n \sim \Bernoulli(P_n)$.  Let $\hat{V}_n$ be the 
matrix of eigenvectors of $A_n$ corresponding to the $K (K + 1) / 2$ most
positive and $K (K - 1) / 2$ most negative eigenvalues. Then 
$\exists \lambda > 0$ and $N \in \mathbb{N}$ such that when $n > N$, 
$\sqrt{n} \hat{V}_n$ obeys the subspace detection property with probability 1.
\end{theorem}
